Universal Anonymous Signatures

### 2022-04-24 prism-backend/research/gsac/app-building-blocks.tex
Proof on non-frameability. Some other minor edits.

\begin{figure}[ht!]
  \begin{minipage}[t]{0.5\textwidth}
    \procedure{$\ExpComBind(\secpar)$}{%
      \Cparm \gets \CSetup(\secpar) \\
      (\msg_0,r_0,\msg_1,r_1) \gets \adv(\Cparm) \\
      \Ccom_0 \gets \CCommit(\Cparm,\msg_0,r_0) \\
      \Ccom_1 \gets \CCommit(\Cparm,\msg_1,r_1) \\
      \pcif \msg_0 \neq \msg_1 \land \Ccom_0 = \Ccom_1: \pcreturn 1 \\
      \pcreturn 0
    }
  \end{minipage}
  \begin{minipage}[t]{0.5\textwidth}
    \procedure{$\ExpComHideb(\secpar)$}{%
      \Cparm \gets \CSetup(\secpar) \\
      (\msg_0,\msg_1,st) \gets \adv(\Cparm) \\
      r \getr \mathcal{R} \\
      \Ccom \gets \CCommit(\Cparm,\msg_b,r) \\
      b' \gets \adv(st,\Ccom) \\
      \pcreturn b'
    }
  \end{minipage}
  \label{fig:nizk-games}
  \caption{Games for commitment schemes.}
\end{figure}

@@

\subsection{Non-Interactive Zero-Knowledge Proofs of Knowledge}
\label{sapp:nizk}

Let \NIZKRel be an NP relation defined by pairs of elements $(\NIZKx,\NIZKw)$,
where \NIZKx is a statement and \NIZKw a witness proving that $(\NIZKx,\NIZKw)
\in \NIZKRel$. For concrete relations, we write $\NIZKRel = \lbrace (\NIZKx),
(\NIZKw): f(x,w) \rbrace$, where $f(x,w)$ is a boolean predicate denoting the
concrete conditions that \NIZKx and \NIZKw need to meet. The set of all $(\NIZKx,
\NIZKw) \in \NIZKRel$ is called the language, or \NIZKLang, for \NIZKRel.
$\NIZKx \notin \NIZKLang$ means that there is no $\NIZKw$ such that $(\NIZKx,
\NIZKw) \in \NIZKRel$. \todo{Check this definition.}

@@

Moreover, a NIZK system has additional algorithms $(\NIZKSimI,\NIZKSimII)$ and
$(\NIZKExtI,\NIZKExtII)$ defined as follows:

\begin{description}
\item[$(\NIZKcrs,\NIZKext) \gets \NIZKSimI(\NIZKsecpar)$.] Produces a \NIZKcrs
  as the \NIZKSetup algorithm, along with a \todo{simulation} trapdoor.
  % The output of \NIZKSimI is required to be indistinguishable from that of
  % \NIZKSetup.
\item[$\NIZKproof \gets \NIZKSimII(\NIZKcrs,\NIZKext,\NIZKx)$.] Given a
  \todo{simulation} trapdoor, and a statement $\NIZKx \in \NIZKLang$,
  produces a simulated proof \NIZKproof of $\NIZKx \in \NIZKLang$.
\item[$(\NIZKcrs,\NIZKext) \gets \NIZKExtI(\NIZKsecpar)$.] Produces a \NIZKcrs
  as the \NIZKSetup algorithm, along with an extraction trapdoor \NIZKext.
\item[$\NIZKw \gets \NIZKExtII(\NIZKcrs,\NIZKext,\NIZKx,\NIZKproof)$.] Given
  an extraction trapdoor \NIZKext produced by \NIZKExtI, and a proof \NIZKproof
  for $\NIZKext \in \NIZKLang$, returns a valid \NIZKw for \NIZKx.
\end{description}

When we want to make explicit the NP relation \NIZKRel to which the previous
algorithms refer to, we use $\NIZKSetup^\NIZKRel,\NIZKProve^\NIZKRel$, 
$\NIZKVerify^\NIZKRel$, etc., and omit the \NIZK prefix and subindex when clear
from context. Altogether, the tuple $(\Setup,\Prove,\Verify,\ExtI,\ExtII,\SimI,
\SimII)$ needs to meet the following properties:

\paragraph{Completeness.} %
Ensures that, for any $(\NIZKx,\NIZKw) \in \NIZKRel)$, any honest prover will be
able to create a proof \NIZKproof that is accepted by any honest verifier, with
overwhelming probability. More precisely, $\Pr\lbrack\ExpNIZKComp\rbrack = 1$
with overwhelming probability, for any p.p.t. \adv, for \ExpNIZKComp in
\figref{fig:nizk-games}.

@@

\paragraph{Knowledge extraction.} %
For all possible adversaries computing a valid proof \NIZKproof over a statement
$\NIZKx \in \NIZKLang$, computed using a \NIZKcrs produced by \ExtI and given
\NIZKext also produced by \ExtI, \ExtII outputs a witness \NIZKw such that
$(\NIZKx,\NIZKw) \in \NIZKRel$. That is, for all valid statements in the
language and a proof, we can extract a valid witness given some special
knowledge ($\NIZKext$). \todo{Capture either here or in the game that the
  output of \ExtI needs to be indistinguishable to the output of \Setup.}

\paragraph{Zero-knowledge.} %
Intuitively, captures that no information can be learned from an statement and
proof pair, beyond their validity (or not). This is captured by requiring the
adversary to distinguish between a run in the real world ($b=0$), where the
setup is done with \Setup, and $\adv$ has access to an honest prover
\Prove; and a run in an ideal world ($b=1$), where the setup is run by
a simulator \SimI, and proofs are generated by \SimII, except when $\adv$
specifies $(\NIZKx,\NIZKw) \notin \NIZKRel$.

\begin{figure}[ht!]
  \begin{minipage}[t]{0.5\textwidth}
    \procedure{$\ExpNIZKComp(\secpar)$}{%
      \NIZKcrs \gets \Setup(\secpar) \\
      (\NIZKx,\NIZKw) \gets \adv(\NIZKcrs) \\
      \pcif (\NIZKx,\NIZKw) \notin \NIZKRel: \pcreturn 0 \\
      \NIZKproof \gets \Prove(\NIZKcrs,\NIZKx,\NIZKw) \\
      b \gets \Verify(\NIZKcrs,\NIZKproof,\NIZKx) \\
      \pcreturn b \\
    }
    \procedure{$\ExpNIZKExt(\secpar)$}{%
      (\NIZKcrs,\NIZKext) \gets \ExtI(\secpar) \\
      (\NIZKx,\NIZKproof) \gets \adv(\NIZKcrs) \\
      \NIZKw \gets \NIZKExtII(\NIZKcrs,\NIZKext,\NIZKx,\NIZKproof) \\
      \pcif \Verify(\NIZKcrs,\NIZKproof,\NIZKx) = 1 \land
      (\NIZKx,\NIZKw) \notin \NIZKRel: \\
      \pcind \pcreturn 0 \\
      \pcreturn 1
    }    
  \end{minipage}
  \begin{minipage}[t]{0.5\textwidth}
    \procedure{$\ExpNIZKSound(\secpar)$}{%
      \NIZKcrs \gets \Setup(\secpar) \\
      (\NIZKproof,\NIZKx) \gets \adv(\NIZKcrs) \\
      \pcif \NIZKx \notin \NIZKLang \land
      \Verify(\NIZKcrs,\NIZKproof,\NIZKx): \\
      \pcind \pcreturn 0 \\
      \pcreturn 1 \\
    }
    \procedure{$\ExpNIZKZKb(\secpar)$}{%
      \pcif b = 0: \\
      \pcind \NIZKcrs \gets \Setup(\secpar) \\
      \pcind \pcreturn \adv^{\Prove(\NIZKcrs,\cdot,\cdot)}(\NIZKcrs) \\
      \pcif b = 1: \\
      \pcind (\NIZKcrs,\NIZKext) \gets \SimI(\secpar) \\
      \pcind \pcreturn \adv^{\Sim'(\NIZKcrs,\NIZKext,\cdot,\cdot)}(\NIZKcrs),~
      \textrm{where} \\
      \pcind \Sim'(\NIZKcrs,\NIZKext,\NIZKx,\NIZKw)~\textrm{returns} \\
      \pcind \pcind \SimII(\NIZKcrs,\NIZKext,\NIZKx)~\pcif (\NIZKx,\NIZKw)
      \in \NIZKRel \\
      \pcind \pcind \bot~\pcif (\NIZKx,\NIZKw) \notin \NIZKRel
    }    
  \end{minipage}
  \label{fig:nizk-games}
  \caption{Games for NIZK schemes.}
\end{figure}

\iffalse
\paragraph{Simulation Soundness.} %
\fi

>>>>

\begin{figure}[ht!]
  \begin{minipage}[t]{0.5\textwidth}
    \procedure{$\ExpComBind(\secpar)$}{%
      \Cparm \gets \CSetup(\secpar) \\
      (\msg_0,r_0,\msg_1,r_1) \gets \adv(\Cparm) \\
      \Ccom_0 \gets \CCommit(\Cparm,\msg_0,r_0) \\
      \Ccom_1 \gets \CCommit(\Cparm,\msg_1,r_1) \\
      \pcif \msg_0 \neq \msg_1 \land \Ccom_0 = \Ccom_1: \pcreturn 1 \\
      \pcreturn 0
    }
  \end{minipage}
  \begin{minipage}[t]{0.5\textwidth}
    \procedure{$\ExpComHideb(\secpar)$}{%
      \Cparm \gets \CSetup(\secpar) \\
      (\msg_0,\msg_1,st) \gets \adv(\Cparm) \\
      r \getr \mathcal{R} \\
      \Ccom \gets \CCommit(\Cparm,\msg_b,r) \\
      b' \gets \adv(st,\Ccom) \\
      \pcreturn b'
    }
  \end{minipage}
  \label{fig:com-games}
  \caption{Games for commitment schemes.}
\end{figure}

@@

\subsection{Simulation-Extractable Non-Interactive Zero-Knowledge
  Proofs of Knowledge}
\label{sapp:nizk}

Let \NIZKRel be an NP relation defined by pairs of elements $(\NIZKx,\NIZKw)$,
where \NIZKx is a statement and \NIZKw a witness proving that $(\NIZKx,\NIZKw)
\in \NIZKRel$. For concrete relations, we write $\NIZKRel = \lbrace (\NIZKx),
(\NIZKw): f(x,w) \rbrace$, where $f(x,w)$ is a boolean predicate denoting the
concrete conditions that \NIZKx and \NIZKw need to meet. The set of all \NIZKx
such that there exists a \NIZKw for which $(\NIZKx,\NIZKw) \in \NIZKRel$ is the
language, or \NIZKLang, for \NIZKRel. $\NIZKx \notin \NIZKLang$ means that
there is no $\NIZKw$ such that $(\NIZKx,\NIZKw) \in \NIZKRel$.

@@

Any zero-knowledge proof of knowledge must meet completeness, soundness and
zero-knowledge,properties. We further need an extra property, called
\emph{simulation-extractability} \cite{cl06}, which amplifies the security
requirements of simulation soundness.
%
To define more formally the properties we need, we have to define three extra
algorithms:

\begin{description}
\item[$(\NIZKcrs,\NIZKtrap) \gets \NIZKSimSetup(\NIZKsecpar)$.] Produces a
  \NIZKcrs as the \NIZKSetup algorithm, along with a trapdoor \NIZKtrap.
\item[$\NIZKproof \gets \NIZKSim(\NIZKcrs,\NIZKtrap,\NIZKx)$.] Given a trapdoor
  \NIZKtrap produced by \NIZKSimSetup, and a statement $\NIZKx \in \NIZKLang$,
  produces a simulated proof \NIZKproof of $\NIZKx \in \NIZKLang$.
\item[$\NIZKw \gets \NIZKExtract(\NIZKcrs,\NIZKtrap,\NIZKx,\NIZKproof)$.] Given
  a trapdoor \NIZKtrap produced by \NIZKSimSetup, and a proof \NIZKproof for
  $\NIZKx \in \NIZKLang$, returns a valid \NIZKw for \NIZKx.
\end{description}

When we want to make explicit the NP relation \NIZKRel to which the previous
algorithms refer to, we use $\NIZKSetup^\NIZKRel,\NIZKProve^\NIZKRel$, 
$\NIZKVerify^\NIZKRel$, etc., and omit the \NIZK prefix and subindex when clear
from context. Altogether, the tuple $(\Setup,\Prove,\Verify,\SimSetup,\Sim,
\Extract)$ needs to meet the following properties:

\paragraph{Completeness.} %
Ensures that, for any $(\NIZKx,\NIZKw) \in \NIZKRel$, any honest prover will be
able to create a proof \NIZKproof that is accepted by any honest verifier, with
overwhelming probability. More precisely, $\Pr\lbrack\ExpNIZKComp\rbrack = 1$
with overwhelming probability, for any p.p.t. \adv, for \ExpNIZKComp in
\figref{fig:nizk-games}.

@@

\paragraph{Zero-knowledge.} %
Intuitively, captures that no information can be learned from a statement and
proof pair, beyond their validity (or not). This is captured by requiring the
adversary to distinguish between a run in the real world ($b=0$), where the
setup is done with \Setup, and $\adv$ has access to an honest prover
\Prove; and a run in an ideal world ($b=1$), where the setup is replaced by
\SimSetup, and proofs are simulated with the help of the trapdoor produced by
\SimSetup, except when $\adv$ specifies $(\NIZKx,\NIZKw) \notin \NIZKRel$. Note
that, in the context of simulation-extractable NIZK, this property not only
requires that the simulated proofs are indistinguishable to the real ones; it
also requires that \SimSetup is indistinguishable from \Setup. All this is
formalised by requiring that $|\Pr[\ExpNIZKZKb(\secpar) = 1 | b = 1] -
\Pr[\ExpNIZKZKb(\secpar) = 1 | b = 0]$ be a negligible function of \secpar,
where \ExpNIZKZKb is as defined in \figref{fig:nizk-games}.

\paragraph{Simulation-Extractability.} As stated, simulation-extractability
is an extension to simulation soundness. In a nutshell,
simulation-extractability requires that, even after having received a polynomial
number of simulated proofs of knowledge, no adversary can output a valid proof
of knowledge from which no witness can be extracted. It implies simulation
soundness, which ``just'' requires that no adversary can produce a valid proof
after having seen polynomially many simulated proofs (but does not guarantee
extraction). Formally, for simulation-extractability we require that
$\Pr[\ExpNIZKSimExt] = 1$ is a negligible function of \secpar, where
\ExpNIZKSimExt is defined in \figref{fig:nizk-games}.

\begin{figure}[ht!]
  \begin{minipage}[t]{0.5\textwidth}
    \procedure{$\ExpNIZKComp(\secpar)$}{%
      \NIZKcrs \gets \Setup(\secpar) \\
      (\NIZKx,\NIZKw) \gets \adv(\NIZKcrs) \\
      \pcif (\NIZKx,\NIZKw) \notin \NIZKRel: \pcreturn 0 \\
      \NIZKproof \gets \Prove(\NIZKcrs,\NIZKx,\NIZKw) \\
      b \gets \Verify(\NIZKcrs,\NIZKproof,\NIZKx) \\
      \pcreturn b \\
    }
    \procedure{$\ExpNIZKSimExt(\secpar)$}{%
      (\NIZKcrs,\NIZKtrap) \gets \SimSetup(\secpar) \\
      (\NIZKx,\NIZKproof) \gets \adv^{\Sim'(\NIZKcrs,\NIZKtrap,\cdot,\cdot)}
      (\NIZKcrs) \\
      \pcind \textrm{Where}~\Sim'(\NIZKcrs,\NIZKtrap,\NIZKx,\NIZKw)~\textrm{returns} \\
      \pcind \pcind
      \Sim(\NIZKcrs,\NIZKtrap,\NIZKx)~\pcif (\NIZKx,\NIZKw) \in \NIZKRel \\
      \pcind \pcind \bot~\pcif (\NIZKx,\NIZKw) \notin \NIZKRel \\
      \NIZKw' \gets \Extract(\NIZKcrs,\NIZKtrap,\NIZKx,\NIZKproof) \\
      \pcif \Verify(\NIZKcrs,\NIZKproof,\NIZKx) = 1 \land
      (\NIZKx,\NIZKw') \notin \NIZKRel~\land \\
      \pcind \NIZKx~\textrm{was not queried to $\Sim$ via $\Sim'$}: \\
      \pcind \pcreturn 1 \\
      \pcreturn 0
    }    
  \end{minipage}
  \begin{minipage}[t]{0.5\textwidth}
    \procedure{$\ExpNIZKSound(\secpar)$}{%
      \NIZKcrs \gets \Setup(\secpar) \\
      (\NIZKproof,\NIZKx) \gets \adv(\NIZKcrs) \\
      \pcif \NIZKx \notin \NIZKLang \land
      \Verify(\NIZKcrs,\NIZKproof,\NIZKx): \\
      \pcind \pcreturn 0 \\
      \pcreturn 1 \\
    }

    \procedure{$\ExpNIZKZKb(\secpar)$}{%
      \pcif b = 0: \\
      \pcind \NIZKcrs \gets \Setup(\secpar) \\
      \pcind \pcreturn \adv^{\Prove(\NIZKcrs,\cdot,\cdot)}(\NIZKcrs) \\
      \pcif b = 1: \\
      \pcind (\NIZKcrs,\NIZKtrap) \gets \SimSetup(\secpar) \\
      \pcind \pcreturn \adv^{\Sim'(\NIZKcrs,\NIZKtrap,\cdot,\cdot)}(\NIZKcrs),~
      \textrm{where} \\
      \pcind \Sim'(\NIZKcrs,\NIZKtrap,\NIZKx,\NIZKw)~\textrm{returns} \\
      \pcind \pcind \Sim(\NIZKcrs,\NIZKtrap,\NIZKx)~\pcif (\NIZKx,\NIZKw)
      \in \NIZKRel \\
      \pcind \pcind \bot~\pcif (\NIZKx,\NIZKw) \notin \NIZKRel
    }    
  \end{minipage}
  \label{fig:nizk-games}
  \caption{Games for Simulation-Extractable NIZK schemes.}
\end{figure}

As studied in \cite{cl06}, simulation-extractable NIZKPoKs formalize the concept
of ``signatures of knowledge'' (see, e.g., \cite{cs97}). Which basically means
that, given an $(\NIZKx,\NIZKw)$ pair from an NP relation, we can treat \NIZKx
as a public key, and \NIZKw as its corresponding private key, and leverage them
to build digital signature schemes -- with the advantage of being able to do so
while proving arbitrary claims, as long as they can be represented as an NP
relation. We note that, given a simulation-extractable NIZK system, it is
straightforward to build a signature of knowledge by adding the message to be
signed in the statement of the NIZK.
